The non-independence of nations and why it matters

### 2022-04-18 appendix.Rmd
pre-submission manuscript edits

\renewcommand{\appendixname}{\bf{Supplementary Material}}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

@@

Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_[@Brownrigg2018]) using the R package _geosphere_[@Hijmans2019]. The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.

Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages. We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0[@Glottolog] into one global language tree (undated and unresolved). We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:

@@

where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$. We then combined these proximities with speaker data from Ethnologue 21[@Ethnologue] and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:

@@

where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$[@Eff2008]. The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.

@@

To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix $D$ based on Euclidean distances between latitude and longitude coordinates. This distance matrix is computed internally by the R package _brms_[@Burkner2017]. The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by _brms_. The model formula is:

@@

>>>>

---
output:
  pdf_document: default
  html_document: default
---
\renewcommand{\appendixname}{\bf{Supplementary Material}}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

@@

Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_; Brownrigg, 2018) using the R package _geosphere_ (Hijmans, 2019). The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.

Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages. We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0 (Hammarström et al., 2017) into one global language tree (undated and unresolved). We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:

@@

where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$. We then combined these proximities with speaker data from Ethnologue 21 (Ethnologue, 2018) and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:

@@

where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$ (Eff, 2008). The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.

@@

To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix $D$ based on Euclidean distances between latitude and longitude coordinates. This distance matrix is computed internally by the R package _brms_ (Bürkner, 2017). The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by _brms_. The model formula is:

@@

Brownrigg, R. _maps: Draw geographical maps._ (2018).

Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. _Journal of Statistical Software_ **80**, 1---28 (2017).

Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. _Structure and Dynamics_ **3**, (2008).

_Ethnologue: Languages of the world._ (SIL International, 2018).

Hammarström, H., Forkel, R., Haspelmath, M. & Bank, S. _Glottolog 3.0._ (Max Planck Institute for the Science of Human History, 2017). doi:[10.5281/zenodo.4061162](https://doi.org/10.5281/zenodo.4061162)

Hijmans, R. J. _geosphere: Spherical trigonometry._ (2019).

### 2022-04-18 manuscript.Rmd
pre-submission manuscript edits

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "2"
    institution   : "Max Planck Institute for the Science of Human History, Jena, Germany"

@@

abstract: |
 Cross-national analyses can illuminate global patternings of national-level outcomes. But since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates. Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence. In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare and have remained so over the past 30 years. When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives. In reanalyses of twelve cross-national relationships, we show that half are no longer significant after controlling for non-independence using global proximity matrices. We urge social scientists to sufficiently control for non-independence in cross-national research.

keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
wordcount         : "5500 words"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
appendix          : "appendix.Rmd"
csl               : nature.csl
urlcolor          : blue
---

```{r setup, include=FALSE}
## !!! for use without tar_make
#library(targets)
#library(tarchetypes)
#library(tidyverse)
#source("R/functionsDAG.R")
#source("R/functionsSignal.R")
#source("R/functionsReview.R")
#source("R/functionsSimulation.R")
#source("R/functionsReplications.R")
#options(tidyverse.quiet = TRUE)
#lapply(c("brms", "cowplot", "conleyreg", "countrycode", 
#         "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
#         "haven", "lmtest", "papaja", "psych", "readxl", 
#         "rstan", "sjlabelled", "tidybayes", "tidyverse"), library, character.only = TRUE)
## !!!
#
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

@@

Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds. At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned 80,000 unique hits. The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations. Such analyses widen the scope of social science beyond Western populations[@Henrich2010;@Pollet2014] and have been used to study, among other topics, the causes of variation in the economic wealth of nations[@Benhabib1994;@Comin2010;@LaPorta1997;@Sachs2001], global patterings of cultural norms and values[@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000], and the universality and diversity of human behaviour and psychology around the world[@Awad2018;@Rhoads2021;@Schulz2019;@Thomson2018].

@@

Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations. This is known as cultural phylogenetic non-independence[@Tylor1889;@Naroll1961;@Naroll1965], and occurs because related nations share cultural traits inherited via descent from a common ancestor. Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original. For example, despite being separated by over 1500 kilometres of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies. More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g. Argentina and Spain).

@@

In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world. To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two dimensions of cultural values from the World Values Survey, traditional vs. secular values and survival vs. self-expression values[@Inglehart2000]. For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \@ref(fig:plotSignal)). In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance. Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results). These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.

@@

Since economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence? To assess this, we systematically searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country". We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis. We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations. For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).

@@

To compare the efficacy of different methods in the literature, we conducted a simulation study. We simulated national-level datasets (n = 236 nations) with varying degrees of spatial or cultural phylogenetic autocorrelation (i.e. non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables. We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects. Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g. [@Currie2009]). In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010]. Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods). To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated proportional to their linguistic proximity[@deVillemereuil2014]. Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e. falsely infering a relationship when none is present).

@@

Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives. This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.

@@

To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation. For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)). For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]). We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g. tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g. female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]). Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).

(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence. (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence. In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates. Regression lines are plotted with 95% confidence intervals.

@@

Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue. In our simulations, controlling for either latitude or longitude did not reduce false positive rates. This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls maybe still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate. High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020]. The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g. continent or language family fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018]. This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g. [@Smith2017]). Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].

Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates. For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes[@Smaldino2016], treat ordinal data as metric[@Liddell2018], incorrectly handle missing values[@Nicholson2017], and ignore best practices in meta-analyses[@Tipton2019]. Why do cross-national studies also rarely account for non-independence? At the institutional level, one possibility is that such practices are incentivised _because_ they inflate false positive rates, increasing the probability of getting published and promoted[@Smaldino2016]. At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations. Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.

@@

Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution. In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95% credible intervals that included zero when controlling for spatial non-independence. Three of these analyses were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes. A recent reanalysis has also called into question various studies of this ilk[@Kelly2020]. We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence. It is interesting that the economic development analyses were more affected by the spatial controls, while the cultural values analyses were more affected by the cultural phylogenetic controls. This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.

We do not wish to dissuade researchers from conducting cross-national studies. On the contrary, such work promises to deepen our understanding of the causes and consequences of economic development and cultural values around the world. Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013]. But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations. Nations are not independent, and our statistical models must reflect this.

@@

Once we had compiled our review database, we calculated the proportion of articles attempting to control for non-independence at least once. We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other. For these proportions, we calculated 95% boostrap confidence intervals with 1000 bootstrap iterations.

@@

Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero. These models were fitted using the _conleyreg_ R package[@Duben2021]. Bayesian models were fitted using the _brms_ R package[@Burkner2017]. Our choice of priors was based on prior predictive simulation. All models converged normally ($\hat{R}$ < 1.1). Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero. We calculated 95% bootstrap confidence intervals for these false positive rates with 1000 bootstrap iterations.

@@

For each individual analysis, we conducted four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance. See Supplementary Methods for full models.

>>>>

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "2"
    institution   : "Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"

@@

abstract: |
 Cross-national analyses test hypotheses about the drivers of global variation in national outcomes. However, since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates. Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence. In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare. When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives. In reanalyses of twelve cross-national relationships, we show that half are no longer significant after controlling for non-independence using global proximity matrices. We urge social scientists to sufficiently control for non-independence in cross-national research.

keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
wordcount         : "5459 words"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
appendix          : "appendix.Rmd"
csl               : nature.csl
urlcolor          : blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

@@

Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national outcomes of all kinds. At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned 80,000 unique hits. The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations. Such analyses widen the scope of social science beyond Western populations[@Henrich2010;@Pollet2014] and have been used to study, among other topics, the causes of variation in the economic wealth of nations[@Benhabib1994;@Comin2010;@LaPorta1997;@Sachs2001], global patterings of cultural norms and values[@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000], and the universality and diversity of human behaviour and psychology around the world[@Awad2018;@Rhoads2021;@Schulz2019;@Thomson2018].

@@

Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations. This is known as cultural phylogenetic non-independence[@Tylor1889;@Naroll1961;@Naroll1965], and occurs because related nations share cultural traits inherited via descent from a common ancestor. Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original. For example, despite being separated by over 1,500 kilometres of ocean, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years. More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g. Argentina and Spain).

@@

In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world. To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two primary dimensions of cultural values from the World Values Survey, traditional vs. secular values and survival vs. self-expression values[@Inglehart2000]. For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \@ref(fig:plotSignal)). In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance. Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results). These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.

@@

Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence? To assess this, we systematically searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country". We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis. We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations. For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).

@@

To compare the efficacy of different methods in the literature, we conducted a simulation study. We simulated national-level datasets (n = 236 nations) with varying degrees of spatial or cultural phylogenetic autocorrelation (i.e. non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables. We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects. Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g. [@Currie2009]). In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010]. Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods). To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014]. Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e. falsely infering a relationship when none is present).

@@

Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are likely to continue to produce false positives. This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.

@@

To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation. For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)). For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]). We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g. tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g. female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]). Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).

(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence. (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence. In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates. Regression lines are plotted with 95% confidence intervals.

@@

Crucially, our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue. In our simulations, controlling for either latitude or longitude did not reduce false positive rates. This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate. High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020]. The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g. continent or language family fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018]. This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g. [@Smith2017]). Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].

Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates. For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants[@Smaldino2016], treat ordinal data as metric[@Liddell2018], incorrectly handle missing values[@Nicholson2017], and ignore best practices in meta-analyses[@Tipton2019]. Why do cross-national studies also rarely account for non-independence? At the institutional level, one possibility is that such practices are incentivised _because_ they generate statistically significant relationships, which increase the probability that a study is published[@Smaldino2016]. At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations. Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.

@@

Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution. In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95% credible intervals that included zero when controlling for spatial non-independence. Three of these analyses were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes. A recent reanalysis has also called into question various studies of this ilk[@Kelly2020]. We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence.

We do not wish to dissuade researchers from conducting cross-national studies. On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values. Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013]. But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations. Nations are not independent, and our statistical models must reflect this.

@@

Once we had compiled our review database, we calculated the proportion of articles attempting to control for non-independence at least once. We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other. For these proportions, we calculated 95% boostrap confidence intervals with 1,000 bootstrap iterations.

@@

Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero. These models were fitted using the _conleyreg_ R package[@Duben2021]. Bayesian models were fitted using the _brms_ R package[@Burkner2017]. Our choice of priors was based on prior predictive simulation. All models converged normally ($\hat{R}$ < 1.1). Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero. We calculated 95% bootstrap confidence intervals for these false positive rates with 1,000 bootstrap iterations.

@@

For each individual analysis, we ran four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance. See Supplementary Methods for full models.
